An upper bound to gas delivery via pressure-swing adsorption in nanoporous materials

### 2020-05-11 gas-adsorption/gas-ads-appendix.tex
made a number of ising edits and a thesis edit.
compute Vopt and add to text. Fix reference to SI in text.

\appendix
\clearpage
\setcounter{table}{0}
\renewcommand{\thetable}{S\arabic{table}}%
\setcounter{figure}{0}
\renewcommand{\thefigure}{S\arabic{figure}}%
\renewcommand{\thesection}{SI \Alph{section}}%

@@

\section{Proof of extremum}
To show that a uniform potential gives the highest deliverable capacity, we consider a potential of interaction between gas and substrate $\V(\vec r)$ that varies with space.  In this proof we will make use of the Fourier transform of this potential:
\begin{align}
    \Vk \equiv \iiint \V(\vec r) e^{-i\vec k\cdot \vec r} d\vec r
\end{align}
The Fourier transform of a uniform potential is a Dirac delta function $\tilde{\V}(\vec k)\propto\delta(\vec k)$. Therefore, to show that a uniform potential extremizes the deliverable capacity, we must show that the functional derivative of the deliverable capacity with respect to $\Vk$ is zero for \emph{nonzero} values of $\vec k$, i.e.
\begin{align}
    \frac{\delta D}{\delta \Vk} &= 0, \text{ if } \vec k\ne 0.
\end{align}
We note that this functional derivative may be non-zero for $\vec k=0$ because we separately maximize with respect to the particular uniform potential $\V$.
This means that
\begin{align}
    \frac{\delta N_H}{\delta \V(\vec k)} &= -\frac{\delta N_L}{\delta \Vk}
\end{align}
where $N_H$ and $N_L$ are the number of particles at the low and high pressure.

>>>>

\setcounter{table}{0}
\renewcommand{\thetable}{S\arabic{table}}%
\setcounter{figure}{0}
\renewcommand{\thefigure}{S\arabic{figure}}%
\renewcommand{\thesection}{SI \Alph{section}}%

@@

\section{Proof of extremum}\label{sec:proof-extremum}
To show that a uniform potential gives the highest deliverable capacity, we consider a potential of interaction between gas and substrate $\V(\vec r)$ that varies with space.  In this proof we will make use of the Fourier transform of this potential:
\begin{align}
    \Vk \equiv \iiint \V(\vec r) e^{-i\vec k\cdot \vec r} d\vec r
\end{align}
The Fourier transform of a uniform potential is a Dirac delta function $\tilde{\V}(\vec k)\propto\delta(\vec k)$. Therefore, to show that a uniform potential extremizes the deliverable capacity, we must show that the functional derivative of the deliverable capacity with respect to $\Vk$ is zero for \emph{nonzero} values of $\vec k$, i.e.
\begin{align}
    \frac{\delta D}{\delta \Vk} &= 0, \text{ if } \vec k\ne 0.
\end{align}
We note that this functional derivative may be non-zero for $\vec k=0$ because we separately maximize with respect to the particular uniform potential $\V$.
This means that
\begin{align}
    \frac{\delta N_H}{\delta \V(\vec k)} &= -\frac{\delta N_L}{\delta \Vk}
\end{align}
where $N_H$ and $N_L$ are the number of particles at the low and high pressure.

### 2020-05-11 gas-adsorption/gas-ads-body.tex
compute Vopt and add to text. Fix reference to SI in text.

The deliverable capacity of gas in our homogeneous substrate is thus:
\begin{align}
    \rho_D(\V,T) &= \rho(\mufull,T) - \rho(\muempty,T),
    \label{eq:DofPhi}
    \\
    &= \rho_g(\mufull-\V,T) - \rho_g(\muempty-\V,T),
\end{align}
where $\mufull$ and $\muempty$ are the chemical potentials corresponding to the pressures $\pfull$ and $\pempty$, respectively.
Thus, $\rho_D$ for our homogeneous substrate as a function of $\V$ is the difference between two shifted versions of $\rho_g(\mu; T)$.
Figure~\ref{fig:methane-298-D} shows the two shifted bulk methane density curves and their difference.
The horizontal axis
is the difference in molar Gibbs free energy at fixed density and temperature between the gas within the substrate and the pure gas (\gst), which is equal to $\V$ in our model.
We see a potential $\V$ that maximizes the deliverable capacity of our ideal substrate, balancing the need to maximize the density at $\pfull$ against the need to minimize residual gas retained at $\pempty$. 
%Our theoretical framework thus recovers an upper bound on the deliverable capacity of methane (298\ K, $\pfull=$65\ bar, $\pempty=$5.8\ bar) in porous materials as 374\ L STP/L.
The optimal deliverable capacity of methane (298\ K, $\pfull=$ 65\ bar, $\pempty=$ 5.8\ bar) in our ideal, spatially uniform substrate is 374\ L STP/L, \red{achieved for $\V_{opt} = ...$}.
%
%As a ballpark, we can expect the attractive potential to be optimal around the maximum slope of $\rho_g(\mu)$.  This is the point where the steric repulsion between the gas atoms starts leading to saturation of the density.

An essential question is whether our idealized substrate offering the spatially uniform potential $\V_{opt}$ for the gas places an upper bound on the deliverable capacity in all rigid porous materials. 
%yields the maximum deliverable capacity over all possible potential energy fields. 
In the SI (see Section~\ref{sec:fixme} \davidsays{Look up section label}), we show that our idealized substrate with spatially uniform potential $\V_{opt}$ yields an \emph{extremum} of the deliverable capacity over all possible \emph{static potential energy fields}, provided the gas does not crystallize at the temperature and in the density range of interest. The potential energy field is \emph{static} if it is unaffected by the presence of the gas; consequently, our model does not apply to flexible materials that undergo gas-induced conformation changes \cite{schneemann2014flexible}. We next argue that this extremum is a maximum by addressing two alternative possibilities: the extremum could turn out to be (i) a saddle point or (ii) a local, rather than global, maximum.

>>>>

The deliverable capacity of gas in our homogeneous substrate is thus:
\begin{align}
    \rho_D(\V,T) &= \rho(\mufull,T) - \rho(\muempty,T),
    \label{eq:DofPhi}
    \\
    &= \rho_g(\mufull-\V,T) - \rho_g(\muempty-\V,T),
\end{align}
where $\mufull$ and $\muempty$ are the chemical potentials corresponding to the pressures $\pfull$ and $\pempty$, respectively.
Thus, $\rho_D$ for our homogeneous substrate as a function of $\V$ is the difference between two shifted versions of $\rho_g(\mu; T)$.
Figure~\ref{fig:methane-298-D} shows the two shifted bulk methane density curves and their difference.
The horizontal axis
is the difference in molar Gibbs free energy at fixed density and temperature between the gas within the substrate and the pure gas (\gst), which is equal to $\V$ in our model.
We see a potential $\V$ that maximizes the deliverable capacity of our ideal substrate, balancing the need to maximize the density at $\pfull$ against the need to minimize residual gas retained at $\pempty$. 
%Our theoretical framework thus recovers an upper bound on the deliverable capacity of methane (298\ K, $\pfull=$65\ bar, $\pempty=$5.8\ bar) in porous materials as 374\ L STP/L.
The optimal deliverable capacity of methane (298\ K, $\pfull=$ 65\ bar, $\pempty=$ 5.8\ bar) in our ideal, spatially uniform substrate is 374\ L STP/L, achieved for $\V_{opt} =$ 5.9\ kJ/mol.
%
%As a ballpark, we can expect the attractive potential to be optimal around the maximum slope of $\rho_g(\mu)$.  This is the point where the steric repulsion between the gas atoms starts leading to saturation of the density.

An essential question is whether our idealized substrate offering the spatially uniform potential $\V_{opt}$ for the gas places an upper bound on the deliverable capacity in all rigid porous materials. 
%yields the maximum deliverable capacity over all possible potential energy fields. 
In the SI (see Section~\ref{sec:proof-extremum}), we show that our idealized substrate with spatially uniform potential $\V_{opt}$ yields an \emph{extremum} of the deliverable capacity over all possible \emph{static potential energy fields}, provided the gas does not crystallize at the temperature and in the density range of interest. The potential energy field is \emph{static} if it is unaffected by the presence of the gas; consequently, our model does not apply to flexible materials that undergo gas-induced conformation changes \cite{schneemann2014flexible}. We next argue that this extremum is a maximum by addressing two alternative possibilities: the extremum could turn out to be (i) a saddle point or (ii) a local, rather than global, maximum.
